\LARGE \bf Energy-Aware Dynamic Planning Algorithm for Autonomous UAVs

### 2020-12-02 icra-2021.tex
fixing the figures
fixes to the text

%\begin{figure*}[t]
%  \centering
%  \input{\figpath/icra-2021_overview2.tikz}
%  \caption{The mission consists of circles and lines covering a polygon. The UAV heads to $\varphi_0$ initially, using the desired velocity vector $\dot{\mathbf{p}}_d$. It later heads to $\varphi_0^-$ reducing the radius $c_{0,1}$ to satisfy the battery constraints. The UAV then converges to $\varphi_1$ in stage $\mathcal{M}_1$, $\varphi_2$ in stage $\mathcal{M}_2$, and so on (the circle $\varphi_2$ is not visible in the figure; it connects $\varphi_1$ and $\varphi_3$).}
%  \label{fig:overview}
%\end{figure*}

@@

Unlike most of the past planning algorithms literature, our algorithm plans both the path and computations. We guide the UAV using a vector field~\cite{de2017guidance} that converges smoothly to a planned trajectory. The use of vector fields for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding}. We denote the trajectory with a mathematical function that we refer to as the trajectory explicit equation (TEE). The path is composed of multiple TEEs. We express the computations with a modeling tool presented in our previous work~\cite{seewald2019coarse}. The algorithm alters the TEE and computations within some parameters using output model predictive control (MPC). These parameters are the maximum, minimum, and current levels of the TEE and computations. The user specifies the parameters in an initial plan with an energy budget. The energy budget is the battery capacity and some other battery parameters. In the remainder of the paper, we adopt the following notation. We refer to the current TEE and computations parameters simply as TEE and computations parameters. We refer to both as parameters. We refer to the constraints sets that delimit the maximum and minimum levels of the parameters as TEE sets and computations sets, and to both as parameters sets. There can be multiple or none computations sets per each task, and the same applies to the TEE. Our goal is to complete the plan with the highest possible parameters as the UAV flies and its batteries drain. The term parameters should not be confused with the battery parameters, which we always mention explicitly. These are fixed values that are not replanned by the algorithm.

@@

The remainder of the paper is organized as follows. The overview of dynamic planning is provided in Section~\ref{sec:prob}, along with a suitable model for the position and energy. The algorithm is proposed in Section~\ref{sec:algo}. Section~\ref{sec:experimental} presents the result and showcases the performances. We propose some conclusions in Section~\ref{sec:conclusion}. Finally, we provide some additional information and examples in the Appendixes~\ref{app:proof-eqv}--\ref{app:simulator}.

@@

We assume that a user-specified initial plan consists of different stages. At each stage, the UAV must follow the TEE within TEE sets and do the computations within the computations sets.

Let us further assume the TEE at stage $i$ can be altered with $\rho$ TEE parameters and the computations with $\sigma$ computations parameters. We then express the TEE as a continuous twice differentiable function $\varphi_i:\mathbb{R}^2\times\mathbb{R}^\rho\rightarrow\mathbb{R}$ of a point and the TEE parameters which returns a metric of the distance between the point and the nominal trajectory. We express the computations as the value of the computations parameters. We discuss the concrete meaning of the value of TEE parameters in Subsection~\ref{sec:model}), and of the computations parameters in Subsection~\ref{sec:computations-model}.

Let us adopt the following mathematical notation. Given an integer $a$, $[a]$ is the set $\{0,1,\dots,a\}$, $[a]^+$ the set $[a]/\{0\}$. $c_{i,j}$ specifues the $j$-th parameter of the $i$-th parameters set $c_i$. $c_{i,j}^0$ is the optimal value of the parameter $c_{i,j}$, and $\underline{c_{i,j}},\overline{c_{i,j}}$ its lower and upper bounds. We store these bounds in the plan specification, which format we discuss in Subsection~\ref{sec:plan}. The set $\langle c_{i,1},c_{i,2},...,c_{i,n}\rangle$ is an ordered list of $n$ parameters.

\begin{defn}[Stage, plan and triggering point]\label{def:mission}
  A \emph{stage} $\Gamma_i$ at time instant $k$ of a plan $\Gamma$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \Gamma_i:=\{\langle&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho}),c_{i,\rho+1},\dots,c_{i,\rho+\sigma}\rangle\\\mid
    &\,\exists\,\,\mathbf{p}_k,\,\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\,\forall j\in[\sigma]^+,\,c_{i,\rho+j}\in\mathbb{S}_{i,j}\,\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the TEE set, and $\mathbb{S}_{i,j}:=[\underline{c}_{i,\rho+j},\overline{c}_{i,\rho+j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th computation set. $\mathbf{p}_k$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$.

@@

\begin{figure}[h]
  \center
  \begin{tikzpicture}[shorten >=1pt,node distance=12ex,on grid,auto] 
    \node        (q_dots0) {$\cdots$};
    \node[state] (q_0) [right=of q_dots0] {$\Gamma_i$};
    \node[state] (q_1) [right=of q_0] {$\Gamma_{i+1}$};
    \node        (q_dots1) [right=of q_1] {$\cdots$};   
    \path[->]
    (q_dots0) edge node{$\mathbf{p}_{\Gamma_{i-1}}$} (q_0)
    (q_0) edge node {$\mathbf{p}_{\Gamma_i}$} (q_1)    
    (q_0) edge [loop above] node {$\mathbf{p}_{k}$} (q_0)
    (q_1) edge node {$\mathbf{p}_{\Gamma_{i+1}}$} (q_dots1)
    (q_1) edge [loop above] node {$\mathbf{p}_{k}$} (q_1)
    ; %end path 
    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=10pt},yshift=0pt]
    (q_dots0.south west) -- (q_dots1.south east) node [black,midway,yshift=-8ex]{$\Gamma$};
  \end{tikzpicture}
  \caption{Finite state machine of the plan}
  \label{fig:state-machine}
\end{figure}

A slice of a finite state machine in Figure~\ref{fig:state-machine} shows the transition between the stages. The triggering point $\mathbf{p}_{i-1}$ allows the transition to stage $\Gamma_i$. The UAV remains in the stage with any generic point $\mathbf{p}_k$. It eventually enters the stage $\Gamma_{i+1}$ with the triggerint point $\mathbf{p}_{\Gamma_i}$ and so on, until it reaches the last triggering point.

The algorithm takes as input the plan and some initial guesses that are necessary for the position and energy estimation. It outputs the new position, the instantaneous energy consumption, and the control input--an action performed evolving the state.

@@

The algorithm reads the plan $\Gamma$ using a specific file format--the plan specification. We refer the reader to Appendix~\ref{app:plan-example} for an example.

The first stage $\Gamma_1$ is expressed in the first line of the file
\begin{algorithmic}[1]
  \State\textbf{\texttt{$k_e$;$r$;$\varphi_1(\mathbf{p}_k,c_{1,1},\dots,c_{1,\rho})$}}\label{code:stage}
\end{algorithmic}
where $k_e$ is a convergence value, $r$ a rotation, and $\varphi_1$ the TEE. We refer the reader to the next subsection where we discuss the physical meaning of the convergence value and rotation. The rotation is expressed in binary format with zero being the counter-clockwise rotation, one the clockwise rotation.

The next line specifies the triggering point $\mathbf{p}_{\Gamma_1}$
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:stage}
  \State\textbf{\texttt{$x$,$y$}}\label{code:trigger-pt}
\end{algorithmic}
where $x$ and $y$ are the two coordinates of the point. The UAV reaches the point within a given radius (a parameter of the algorithm) and enters the next stage $\Gamma_2$. The coordinates can be expressed in function of the TEE parameters $c_{1,1},\dots,c_{1,\rho}$. Generally, one can express the coordinates in function of the $i$-th TEE parameters $c_{i,1},\dots,c_{i,\rho}$, or any previous TEE parameters, propagating the information therein if necessary. See Figure~\ref{fig:state-machine2} where we neglected the $\sigma=0$ computations parameters for simplicity $c_i=c_{i,1},\dots,c_{i,\rho}$.

\begin{figure}[h]
  \center
  \begin{tikzpicture}[shorten >=1pt,node distance=24ex,on grid,auto] 
    \node        (q_dots0) {$\cdots$};
    \node[state] (q_0) [right=of q_dots0] {$\langle\varphi_i,c_i\rangle$};
    \node        (q_dots1) [right=of q_0] {$\cdots$};   
    \path[->]
    (q_dots0) edge node{$\mathbf{p}_{\Gamma_{i-1}}(c_1,\dots,c_{i-1})$} (q_0)
    (q_0) edge node {$\mathbf{p}_{\Gamma_i}(c_1,\dots,c_{i})$} (q_dots1)    
    (q_0) edge [loop above] node {$\mathbf{p}_{k}$} (q_0)
    ; %end path
  \end{tikzpicture}
  \caption{Detail of the stage $\Gamma_i$ in the finite state machine}
  \label{fig:state-machine2}
  \end{figure}

@@

The following $\rho+\sigma$ lines
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:trigger-pt}
  \State\textbf{\texttt{[$\underline{c}_{1,j}$,$\overline{c}_{1,j}$]}}\label{code:parameter-set}
\end{algorithmic}
specifies the lower $\underline{c}_{1,j}$ and upper $\overline{c}_{1,j}$ bound of the $j$-th parameter set. They are in ascending order and the TEE parameters sets are specified before the computations parameters sets.

@@

Let us assume that the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using the TEE parameters $c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:overview}). $\varphi_i$ is a function of the two coordinates and is equal to zero when a point $\mathbf{p}_k$ is on the TEE. Physically, this means the UAV is flying exactly over the nominal trajectory. The TEE parameters allows to change the TEE. They are a way to alter the nominal trajectory in the initial plan.
In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to alter the optimal TEE parameters $c_i^0$ (here we again neglect the $\sigma=0$ computations parameters for simplicity), such that $\varphi_i(\mathbf{p}_k,c_i^0)$ has the highest instantaneous energy consumption (while still respecting the energy budget).

We derive the new position $\mathbf{p}_{k+1}$ using the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, computing its vector field $\nabla\varPhi:=(\partial\varPhi/\partial x,\partial\varPhi/\partial y)$, and the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the rotation (which influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:overview}).

@@

We model the energy using some energy coefficients $\mathbf{q}\in\mathbb{R}^m$ that characterize the energy signal. We refer to the instantaneous energy consumption evolution simply as energy signal. The coefficients are derived from Fourier analysis (the size of the energy coefficients vector $m$ is related to the order of a Fourier series). Here we derive the energy model. Later we assess the energy contribution of the computations in Subsection~\ref{sec:computations-model}. This approach is adapted from our earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing UAV~\cite{seewald2020mechanical}.

We show a relation between the energy signal and the energy coefficients in Theorem~\ref{thm:state-vs-energy}. We show after the main results how this approach allows us variability in terms of the systems behaving periodically, piece-wise periodically, or merely linearly with sporadic periodicity.

Let us consider a periodic signal of period $T\in\mathbb{Z}_{> 0}$, and a Fourier series of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of modeling of the signal
\begin{equation}\label{eq:fourier}
  h(t)=a_0/T+(2/T)\sum_{j=1}^{r}{\left(a_j\cos{\omega jt}+b_j\sin{\omega jt}\right)},
\end{equation}
where $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ maps time to the signal, $\omega:=2\pi/T$ is the angular frequency, and $a,b\in\mathbb{R}$ the Fourier series coefficients.

@@

Let us consider a linear time-invariant (LTI) state-space model
\begin{equation}\label{eq:state-perf}\begin{split}
  \dot{\mathbf{q}}&=A\mathbf{q}+B\mathbf{u}+\mathbf{w},\\
  y&=C\mathbf{q}+v,
\end{split}\end{equation}
where $y\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ are the energy coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    0&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_j\begin{bmatrix}0 & \omega j \\ -\omega j & 0\end{bmatrix},\\
  C&=(1/T)\left[\begin{array}{cccccc}
    1 & 1 & 0 &\cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}\in\mathbb{R}^m$ with $m=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the top left entry is one, the diagonal entries are $A_1,\dots,A_r$, the remaining entries are zeros.

\begin{lem}[Signals equality]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are zeros) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then Equation~(\ref{eq:fourier}) describes the same signal as Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the equality of the two signals is achieved by a proper choice of items of matrices $A,C$ and the initial guess $\mathbf{q}_0$. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof of the equality where we justify the choices of the items of the matrices and of the initial guess. 
\end{proof}

@@

Let us suppose that at time instant $k$ the plan reached $i$-th stage $\Gamma_i$. The control along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}
  \mathbf{u}&=\begin{bmatrix}c_{i,1} & \cdots & c_{i,\rho} & c_{i,\rho+1} & \cdots & c_{i,\rho+\sigma}\end{bmatrix}^T,\\
  B\mathbf{u}&:=B_i(\mathbf{u}_k-\mathbf{u}_{k-1}),\,\,\,B_i=\left[\begin{array}{c} \nu_i\\\mathbf{0}\end{array}\right],
\end{split}\end{equation}
where $\mathbf{u}\in\mathbb{R}^n$ is the control with $n=\rho+\sigma$, $B\in\mathbb{R}^{m\times n}$, and $\mathbf{u}_{-1}=\mathbf{0}$. Moreover, the top row of $B$ contains gain factors $\nu_i=\begin{bmatrix}\nu_{i,1} & \cdots & \nu_{i,\rho} & \nu_{i,\rho + 1}& \cdots & \nu_{i,\rho+\sigma}\end{bmatrix}$, quantifying the contribution of a given parameter to the instantaneous energy consumption. The other entries of $B$ are zeros.

The first $\rho$ gain factors quantifies the contribution of the $i$-th TEE parameters. A first guess of these parameters is derived empirically. They are later refined by the algorithm. The following $\sigma$ gain factors quantifies the contribution of the $i$-th computations parameters. They are derived using the modeling tool. We use the gain factors to select the optimal control $\mathbf{u}^0$.

Equation~(\ref{eq:state-control}) accounts for the energy due to the change of parameters $\mathbf{u}_k-\mathbf{u}_{k-1}$. For instance, when the TEE $\varphi_0$ is a circle (see Figure~\ref{fig:overview}), a decrement in the TEE parameter $c_{0,1}$--the radius of the circle--adds a negative contribution. It thus simulates the lowering of instantaneous energy consumption $\nu_{0,1}(c_{0,1}-c_{0,1}^-)<0$.

>>>>

@@

Unlike most of the past planning algorithms literature, our algorithm plans both the path and computations. We guide the UAV using a vector field~\cite{de2017guidance} that converges smoothly to a planned trajectory. The use of vector fields for planning is widely discussed in the literature~\cite{lindemann2005smoothly,gonccalves2010vector,panagou2014motion,zhou2014vector,kapitanyuk2017guiding,de2017guidance}. We denote the trajectory with a mathematical function that we refer to as the trajectory explicit equation (TEE). The path is composed of multiple TEEs. We express the computations with a modeling tool presented in our previous work~\cite{seewald2019coarse}. The algorithm alters the TEE and computations within some parameters using output model predictive control (MPC). These parameters are the maximum, minimum, and current levels of the TEE and computations. The user specifies the parameters in an initial plan with an energy budget. The energy budget is the battery capacity and some other battery parameters. In the remainder of the paper, we adopt the following notation. We refer to the current TEE and computations parameters simply as TEE and computations parameters. We refer to both as parameters. We refer to the constraints sets that delimit the maximum and minimum levels of the parameters as TEE set and computations sets, and to both as parameters sets. There can be multiple or none computations parameters per each task, and the same applies to the TEE. There is one TEE set but multiple computations sets, one per each computation parameter. Our goal is to complete the plan with the highest possible parameters as the UAV flies and its batteries drain. The term parameters should not be confused with the battery parameters, which we always mention explicitly. These are fixed values that are not replanned by the algorithm.

@@

The remainder of the paper is organized as follows. The overview of dynamic planning is provided in Section~\ref{sec:prob}, along with a suitable model for the position and energy. The algorithm is proposed in Section~\ref{sec:algo}. Section~\ref{sec:experimental} presents the result and showcases the performances. We propose some conclusions in Section~\ref{sec:conclusion}. Finally, we provide some additional information and examples in Appendixes~\ref{app:proof-eqv}--\ref{app:simulator}.

@@

The algorithm takes as input a user-specified initial plan that consists of different stages and some initial guesses that are necessary for the position and energy estimation. At each stage, the UAV must follow the TEE within TEE set and do the computations within the computations sets. The algorithm outputs the new position, the instantaneous energy consumption, and the control.

Let us assume the TEE at stage $i$ can be altered with $\rho$ TEE parameters and the computations with $\sigma$ computations parameters. We then express the TEE as a continuous twice differentiable function $\varphi_i:\mathbb{R}^2\times\mathbb{R}^\rho\rightarrow\mathbb{R}$ of a point and the TEE parameters. The TEE returns a metric of the distance between the point and the nominal trajectory. We express the computations as the value of the computations parameters. We discuss the concrete meaning of the value of TEE parameters in Subsection~\ref{sec:model}, and of the computations parameters in Subsection~\ref{sec:computations-model}.

Let us adopt the following mathematical notation. Given an integer $a$, $[a]$ is the set $\{0,1,\dots,a\}$, $[a]^+$ the set $[a]/\{0\}$. $c_{i,j}$ specifues the $j$-th parameter of the $i$-th parameters set $c_i$. $c_{i,j}^0$ is the optimal value of the parameter $c_{i,j}$, and $\underline{c}_{i,j},\overline{c}_{i,j}$ its lower and upper bounds. We store these bounds in the plan specification, which format we discuss in Subsection~\ref{sec:plan}. The set $\langle c_{i,1},c_{i,2},...,c_{i,n}\rangle$ is an ordered list of $n$ parameters.

\begin{defn}[Stage, plan, and triggering point]\label{def:mission}
  A \emph{stage} $\Gamma_i$ at time instant $k$ of a plan $\Gamma$ is defined as the ordered list
  \begin{equation}\label{eq:mission}\begin{split}
    \Gamma_i:=\{\langle&\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho}),c_{i,\rho+1},\dots,c_{i,\rho+\sigma}\rangle\\\mid
    &\,\exists\,\,\mathbf{p}_k,\,\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i,\,\\
      &\,\forall j\in[\sigma]^+,\,c_{i,\rho+j}\in\mathbb{S}_{i,j}\,\},
  \end{split}\end{equation}
  where $\mathbb{C}_i:=[\underline{c}_i,\overline{c}_i]\subseteq\mathbb{R}$ is the TEE set, and $\mathbb{S}_{i,j}:=[\underline{c}_{i,\rho+j},\overline{c}_{i,\rho+j}]\subseteq\mathbb{Z}_{\geq 0}$ the $j$-th computation set. $\mathbf{p}_k$ is a point of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$.

@@

\begin{figure}[h]
  \center
  \begin{tikzpicture}[shorten >=1pt,node distance=12ex,on grid,auto] 
    \node        (q_dots0) {$\cdots$};
    \node[state] (q_0) [right=of q_dots0] {$\Gamma_i$};
    \node[state] (q_1) [right=of q_0] {$\Gamma_{i+1}$};
    \node        (q_dots1) [right=of q_1] {$\cdots$};   
    \path[->]
    (q_dots0) edge node{$\mathbf{p}_{\Gamma_{i-1}}$} (q_0)
    (q_0) edge node {$\mathbf{p}_{\Gamma_i}$} (q_1)    
    (q_0) edge [loop above] node {$\mathbf{p}_{k}$} (q_0)
    (q_1) edge node {$\mathbf{p}_{\Gamma_{i+1}}$} (q_dots1)
    (q_1) edge [loop above] node {$\mathbf{p}_{k}$} (q_1)
    ; %end path 
    \draw [decorate,decoration={brace,amplitude=10pt,mirror,raise=10pt},yshift=0pt]
    (q_dots0.south west) -- (q_dots1.south east) node [black,midway,yshift=-8ex]{$\Gamma$};
  \end{tikzpicture}
  \caption{The plan defined as a finite state machine}
  \label{fig:state-machine}
\end{figure}

A slice of the plan in Figure~\ref{fig:state-machine} shows the transition between the stages. The triggering point $\mathbf{p}_{\Gamma_{i-1}}$ allows the transition to stage $\Gamma_i$. The UAV remains in the stage with any generic point $\mathbf{p}_k$. It eventually enters the stage $\Gamma_{i+1}$ with the triggerint point $\mathbf{p}_{\Gamma_i}$ and so on, until it reaches the last triggering point.

@@

The algorithm reads the plan $\Gamma$ from a file--the plan specification--with a specific file format. We refer the reader to Appendix~\ref{app:plan-example} for an example.

The first stage $\Gamma_1$ is expressed in the first line of the file
\begin{algorithmic}[1]
  \State\textbf{\texttt{$k_e$;$r$;$\varphi_1(\mathbf{p}_k,c_{1,1},\dots,c_{1,\rho})$}}\label{code:stage}
\end{algorithmic}
where $k_e$ is a convergence value, $r$ a rotation, and $\varphi_1$ the TEE. In the next subsection where we discuss the physical meaning of the convergence value and rotation. The rotation is expressed in binary format with zero being the counter-clockwise rotation, one the clockwise rotation.

The next line specifies the triggering point $\mathbf{p}_{\Gamma_1}$
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:stage}
  \State\textbf{\texttt{$x$,$y$}}\label{code:trigger-pt}
\end{algorithmic}
where $x$ and $y$ are the two coordinates of the point. The UAV reaches the point within a given radius (a parameter of the algorithm) and enters the next stage $\Gamma_2$. The coordinates can be expressed in function of the TEE parameters $c_{1,1},\dots,c_{1,\rho}$. Generally, one can express the coordinates in function of the $i$-th TEE parameters $c_{i,1},\dots,c_{i,\rho}$, or any previous TEE parameters, propagating the information therein if necessary. See Figure~\ref{fig:state-machine2} where, for simplicity, we neglected the $\sigma=0$ computations parameters $c_i=c_{i,1},\dots,c_{i,\rho}$.

\begin{figure}[h]
  \center
  \begin{tikzpicture}[shorten >=1pt,node distance=23ex,on grid,auto] 
    \node        (q_dots0) {$\cdots$};
    \node[state] (q_0) [right=of q_dots0] {$\Gamma_i$};
    \node        (q_dots1) [right=of q_0] {$\cdots$};   
    \path[->]
    (q_dots0) edge node{$\mathbf{p}_{\Gamma_{i-1}}(c_1,\dots,c_{i-1})$} (q_0)
    (q_0) edge node {$\mathbf{p}_{\Gamma_i}(c_1,\dots,c_{i})$} (q_dots1)    
    (q_0) edge [loop above] node {$\mathbf{p}_{k}$} (q_0)
    ; %end path
  \end{tikzpicture}
  \caption{Detail of the stage $\Gamma_i$ in the finite state machine}
  \label{fig:state-machine2}
  \end{figure}

@@

The line
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:trigger-pt}
  \State\textbf{\texttt{[$\underline{c}_1$,$\overline{c}_1$]}}\label{code:tee-set}
\end{algorithmic}
specifies the lower $\underline{c}_{1}$ and upper $\overline{c}_{1}$ bounds of the TEE $\varphi_1$.

The following $\sigma$ lines
\begin{algorithmic}[1]
  \setcounterref{ALG@line}{code:tee-set}
  \State\textbf{\texttt{[$\underline{c}_{1,j}$,$\overline{c}_{1,j}$]}}\label{code:parameter-set}
\end{algorithmic}
specifies the lower $\underline{c}_{1,j}$ and upper $\overline{c}_{1,j}$ bounds of the $j$-th computation parameter set. They are in ascending order.

This means that for the first stage, the algorithm can alter the TEE satisfying $\varphi_1\in\mathbb{C}_1$, and it can alter the computations individually. In fact the $j$-th computation parameter has to be in $\mathbb{S}_{1,j}$.

@@

Let us assume that the set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using the TEE parameters $c_{i,1},...,c_{i,\rho}$ (the gray area in Figure~\ref{fig:overview}). $\varphi_i$ is a function of the two coordinates and is equal to zero when a point $\mathbf{p}_k$ is on the TEE. Physically, this means the UAV is flying exactly over the nominal trajectory. The TEE parameters allows to change the TEE. They are a way to alter the nominal trajectory in the initial plan.
In fact, the algorithm uses the set from Equation~(\ref{eq:area}) to alter the optimal TEE parameters $c_i^0$ (here we again neglect for simplicity the $\sigma=0$ computations parameters), such that $\varphi_i(\mathbf{p}_k,c_i^0)$ has the highest instantaneous energy consumption (while still respecting the energy budget).

We derive the new position $\mathbf{p}_{k+1}$ using the function $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, computing its vector field $\nabla\varPhi:=\begin{bmatrix}\partial\varPhi/\partial x & \partial\varPhi/\partial y\end{bmatrix}^T$, and the direction to follow in the form of velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $E$ specifies the rotation (which influence the tracking direction), and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind ($w\in\mathbb{R}$ in Figure~\ref{fig:overview}).

\begin{figure}[h]
  \centering
  \input{\figpath/path2.tikz}
  \caption{Example of the alteration of a TEE parameter}
  \label{fig:overview}
\end{figure}

@@

We model the energy using some energy coefficients $\mathbf{q}\in\mathbb{R}^m$ that characterize the energy signal. We refer to the instantaneous energy consumption evolution simply as the energy signal. The coefficients are derived from Fourier analysis (the size of the energy coefficients vector $m$ is related to the order of a Fourier series). We enhance the energy model with the energy contribution of the computations in Subsection~\ref{sec:computations-model}.

We proof a relation between the energy signal and the energy coefficients in Theorem~\ref{thm:state-vs-energy}. We show after the main results how this approach allows us variability in terms of the systems behaving periodically, piece-wise periodically, or merely linearly with sporadic periodicity.

Let us consider a periodic signal of period $T\in\mathbb{Z}_{> 0}$, and a Fourier series of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ for the purpose of modeling of the signal
\begin{equation}\label{eq:fourier}
  h(t)=a_0/T+(2/T)\sum_{j=1}^{r}{\left(a_j\cos{\omega jt}+b_j\sin{\omega jt}\right)},
\end{equation}
where $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ maps time to the instantaneous energy consumption, $\omega:=2\pi/T$ is the angular frequency, and $a,b\in\mathbb{R}$ the Fourier series coefficients.

@@

Let us consider a linear time-invariant (LTI) state-space model
\begin{equation}\label{eq:state-perf}\begin{split}
  \dot{\mathbf{q}}&=A\mathbf{q}+B\mathbf{u}+\mathbf{w},\\
  y&=C\mathbf{q}+v,
\end{split}\end{equation}
where $y\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption, and $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are the contributions of the noice.

The state $\mathbf{q}$ are the energy coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    0&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_j\begin{bmatrix}0 & \omega j \\ -\omega j & 0\end{bmatrix},\\
  C&=(1/T)\left[\begin{array}{cccccc}
    1 & 1 & 0 &\cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}\in\mathbb{R}^m$ with $m=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the top left entry is zero, the diagonal entries are $A_1,\dots,A_r$, the remaining entries are zeros.

\begin{lem}[Signals equality]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w},v$ are zeros) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then, the signal $h$ in Equation~(\ref{eq:fourier}) is equal to the signal $y$ in Equation~(\ref{eq:state-perf}).
\end{lem}
\begin{proof}
The equivalence of the models is trivial and the equality of the two signals is achieved by a proper choice of items of matrices $A,C$ and the initial guess $\mathbf{q}_0$. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof of the equality where we justify the choices of the items of the matrices and of the initial guess. 
\end{proof}

@@

Let us suppose that at time instant $k$ the plan reached the $i$-th stage $\Gamma_i$.

The control along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}
  \mathbf{u}&=\begin{bmatrix}c_{i,1} & \cdots & c_{i,\rho} & c_{i,\rho+1} & \cdots & c_{i,\rho+\sigma}\end{bmatrix}^T,\\
  B\mathbf{u}&:=B_i(\mathbf{u}_k-\mathbf{u}_{k-1}),\,\,\,B_i=\left[\begin{array}{c} \nu_i\\\mathbf{0}\end{array}\right],
\end{split}\end{equation}
where $\mathbf{u}\in\mathbb{R}^n$ is the control with $n=\rho+\sigma$, $B\in\mathbb{R}^{m\times n}$, and $\mathbf{u}_{-1}=\mathbf{0}$. Moreover, the top row of $B$ contains gain factors $\nu_i=\begin{bmatrix}\nu_{i,1} & \cdots & \nu_{i,\rho} & \nu_{i,\rho + 1}& \cdots & \nu_{i,\rho+\sigma}\end{bmatrix}$, quantifying the contribution of a given parameter to the instantaneous energy consumption. The other entries of $B$ are zeros.

The first $\rho$ gain factors quantifies the contribution of the $i$-th TEE parameters. A first guess of these parameters is derived empirically. They are later refined by the algorithm. The following $\sigma$ gain factors quantifies the contribution of the computations parameters. They are derived using the modeling tool. We use the gain factors to select the optimal control $\mathbf{u}^0$.

Equation~(\ref{eq:state-control}) accounts for the energy due to the change of parameters $\mathbf{u}_k-\mathbf{u}_{k-1}$. For instance, when the TEE $\varphi_1$ is a circle (see Figure~\ref{fig:overview}), a decrement in the TEE parameter $c_{1,1}$--the radius of the circle--adds a negative contribution. It thus simulates the lowering of instantaneous energy consumption $\nu_{1,1}(c_{1,1}-c_{1,1}^-)<0$.
